---
title: "Reproducible Research Week-2 Assignment"
author: "Ashutosh"
date: "17/02/2020"
output:
  html_document: default
  pdf_document: default
---
#Reproducible Research Week-2 Assignment

## Loading and preprocessing the data
```{r}
activitydata <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
library(ggplot2)
total.steps <- tapply(activitydata$steps, activitydata$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```
## What is the average daily activity pattern?
```{r}
library(ggplot2)
averages <- aggregate(x=list(steps=activitydata$steps), by=list(interval=activitydata$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken")
```
the maximum number of steps in 5-minute interval

```{r}
averages[which.max(averages$steps),]
```

## Imputing missing values
```{r}
missingvalues <- is.na(activitydata$steps)
# How many missing
table(missingvalues)
```

# Replace each missing value with the mean value of its 5-minute interval
```{r}
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}
filled.data <- activitydata
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```

Histogram of the total number of steps taken each day and mean and median of total number of steps.

```{r}
total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps)
median(total.steps)
```

Mean and median values are higher after imputing missing data because in the original data, there are some days with `steps` values `NA` for any `interval`. The total number of steps taken in such days are set to 0s by default. Therefore after replacing missing `steps` values with the mean `steps`
of associated `interval` value, these 0 values are removed from the histogram
of total number of steps taken each day.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN=weekday.or.weekend)
```

Now plots of average number of steps taken on weekdays and weekends.
```{r}
averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
    xlab("5-minute interval") + ylab("Number of steps")
